---
title: 'IDA 2019: Homework 2'
author: "Group 01"
date: "11/18/2019"
output: html_document
---
# Data wrangling & summary statistics

```{r echo=FALSE}

knitr::opts_chunk$set(
  warning = FALSE, # supress warnings per default 
  message = FALSE  # supress messages per default 
)

```

```{r echo=FALSE}

library(tidyverse)

```

## Exercise 1: Fictitious data from a button-press reaction time experiment
### Exercise 1.A: Tidy up the mess

```{r}

messy_data <- tribble(
  ~subject_id,  ~choices,  ~reaction_times,
  1,            "A,B,B",   "312 433 365",
  2,            "B,A,B",   "393 491 327",
  3,            "B,A,A",   "356 313 475",
  4,            "A,B,B",   "292 352 378"
)

```

```{r}

tidy_data_choice <- messy_data %>% separate(
  col = choices,
  into = str_c("C_", 1:3),
  sep = ",",
  convert = TRUE
) %>% select(subject_id, C_1, C_2, C_3) %>% pivot_longer(
  cols = -1,
  names_to = "condition",
  values_to = "response"
)
tidy_data_choice

tidy_data_RT <- messy_data %>% separate(
  col = reaction_times,
  into = str_c("C_", 1:3),
  sep = " ",
  convert = TRUE
) %>% select(-2) %>% pivot_longer(
  cols = -1,
  names_to = "condition",
  values_to = "RT"
)
tidy_data_RT

tidy_data <- full_join(tidy_data_choice, tidy_data_RT)
tidy_data

```

### Exercise 1.B: Summarize the reaction times

```{r}

tidy_data %>% 
  group_by(condition) %>% 
  summarise(
    mean_RT = mean(RT)
  )

```

```{r}

tidy_data %>% 
  group_by(subject_id) %>% 
  summarise(
    mean_RT = round(mean(RT))
  )

```

## Exercise 2: The King of France visits IDA
### Exercise 2.A: Experimental design

1) 
Yes, it is a factorial design. There is one factor: "Condition" with 7 levels, namely "filler", "background_check" and the 5 critical conditions.

2)
Within-subjects design, since every participant contributes at least one observation to each experimental condition.

3) Pro: You need fewer participants, Con: There is possible cross-contamination between conditions.

4) It is a repeated-measures design, because we have 5 critical conditions for each vignette.

5) The dependent variable of the experiment is "condition". The variable type is an ordered factor.

### Exercise 2.B: Exploring IDA’s King of France

```{r}
data_KoF_raw_IDA <- 
  read_csv(url('https://raw.githubusercontent.com/michael-franke/intro-data-analysis/master/data_sets/king-of-france_data_raw_IDA.csv'))

glimpse(data_KoF_raw_IDA)
```

```{r}
nrow(data_KoF_raw_IDA)
data_KoF_raw_IDA %>% pull(submission_id) %>% unique %>% length
data_KoF_raw_IDA %>% pull(comments) %>% unique
data_KoF_raw_IDA %>% pull(languages) %>% unique
data_KoF_raw_IDA %>% pull(age) %>% mean(na.rm = TRUE)
data_KoF_raw_IDA %>% pull(age) %>% summary
```
7.
- submission_id: ordinal
- RT: metric
- correct: boolean
- education: nominal
- item_version: nominal
- question: nominal
- response: boolean
- timeSpent: metric
- trial_name: nominal
- trial_number: ordinal
- trial_type: nominal
- vignette: nominal

## Preprocessing the data
### Selecting and creating the relevant columns

```{r}

data_KoF_processed_IDA <- data_KoF_raw_IDA %>% 
  # discard practice trials
  filter(trial_type != "practice") %>% 
  mutate(
    # add a 'condition' variable
    condition = case_when(
      trial_type == "special" ~ "background check",
      trial_type == "main" ~ str_c("Condition ", item_version),
      TRUE ~ "filler"
    ) %>% 
      factor( 
        ordered = T,
        levels = c(str_c("Condition ", c(0, 1, 6, 9, 10)), "background check", "filler")
      )
  ) %>% select(submission_id, trial_number, condition, vignette, question, correct, response)

```

### Tidy?
Yes, this last data representation is tidy since each variable from our case study forms a column and each observation forms a line. Also the third condition is met for all of the filled cells form a table. Also, there is no redundant variable in the data representation.

## Towards testing a hypothesis

```{r}
data_KoF_processed_IDA %>% 
  filter(
    condition %in% c("Condition 0", "Condition 1")
    ) %>% group_by(condition) %>% summarise(proportion_true = mean(response))
```