---
title: "IDA 2019: Homework 1"
author: "Group 01"
date: "November 11, 2019"
output: html_document
---

# Task 1
## The data set I: Creating a table in RMarkdown
### Overview of variables in the data set

**Variable**   | **Variable type** | **Values**
:------------: | :---------------: | :-----------------------
brain regions  | nominal           | (Rostral middle frontal, Precentral, Lateral Occipital,  Transverse temporal, Temporal pole)
surface area   | metric            | (5941.8, 4718.8, 4672.9, 799.48, 443.3)
thickness      | metric            | (2.59, 2.74, 2.3, 2.52, 3.66)
hemisphere     | binary            | (L, L, L, R, R)

# Task 2
## Loading relevant packages

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

# Task 3
## The data set II: Creating a tibble in R

```{r}
brain_data <- tibble(
              brain_regions = factor(c("Rostral middle frontal","Precentral","Lateral Occipital","Transverse temporal","Temporal pole")),
              surface_area = c(5941.8, 4718.8, 4672.9, 799.48, 443.3), 
              thickness = c(2.59, 2.74, 2.3, 2.52, 3.66), 
              hemisphere = c("L", "L", "L", "R", "R"), 
              .rows=5
              )
brain_data
```

# Task 4

```{r}
brain_data$brain_regions <- fct_recode(brain_data$brain_regions,  "Banks superior temporal" = "Transverse temporal")
brain_data
```

# Task 5

```{r}
precentral_thickness <- brain_data$thickness[2]
precentral_thickness
```

# Task 6
## The data set III: Creating a tibble in R

```{r}
brain_data2 <- tibble(
               brain_regions = factor(c("Rostral middle frontal","Precentral","Lateral Occipital","Transverse temporal","Temporal pole")),
               surface_area = c(5941.8, 4718.8, 4672.9, 799.48, 443.3), 
               thickness = c(2.59, 2.74, 2.3, 2.52, 3.66), 
               hemisphere = c("L", "L", "L", "R", "R"),
               volume = surface_area * thickness,
               .rows=5
               )
brain_data2
```

# Task 7
Usual data frames do not allow dynamic construction whereas tibbles do. That means that tibbles allow you to immediately work with variables in the same function call in which they were defined.

# Task 8
```{r}
regions_v1 <- brain_data2$brain_regions
regions_v1
```

# Task 9
```{r}
typeof(regions_v1)
```

# Task 10
```{r}
length(regions_v1)
```

# Task 11
```{r}
brain_data2$hemisphere %>% unique %>% length
```

# Task 12
Returns only brain regions and only those in right hemisphere.

# Task 13
```{r}
x <- brain_data2$surface_area
y <- brain_data2$thickness

volume_calc <- function(x,y) {
  return(x * y)
}

volume_compare <- list(volume1 = brain_data2$volume, volume2 = volume_calc(x,y))
volume_compare
```

# Task 14
```{r}
volume_true <- c(17439, 14351, 12150, 2386, 2280.1)
volume_calculated <- unlist(volume_compare[1])

volume_diff <- function(x,y) {
  return(x-y)
}
```

# Task 15
```{r}
brain_data3 <- tibble(
               brain_regions = regions_v1,
               calculated_volume = volume_diff(volume_calculated, volume_true)
               )
brain_data3
```

# Task 16
Which brain region has the biggest difference between calculated and measured volume?

# Task 17
```{r}
family <- c("Gomez", "Morticia", "Pugsley", "Wednesday", "Uncle Fester", "Grandma")
map_chr(family,
        function(i) {
          if (i == "Uncle Fester" | i == "Grandma" ) {
            return(i)
          }
          else {
            return(str_c(i, " Adams"))
          }
        }  
)
```